Early Observations of the Type Ia Supernova \abc:\\ \textbf{a Case of Interaction with Nearby, Unbound Material and/or} Strong Ejecta Mixing

### 2017-11-17 doc/ms.tex
update text in the conclusions

\begin{itemize}
    \item Extrapolation of the early light curve shows that the initial detection of \abc\ occurred only $0.15\pm_{0.07}^{0.15}$~d after the time of first light, $t_0$.

@@

    \item Assuming that $v_\mathrm{ph} \propto t^{-0.22}$, then $t_\mathrm{exp} \approx t_0$. A short dark phase, as this implies, is likely the result of either strong $^{56}$Ni mixing or interaction of the SN ejecta with diffuse material.

@@

    \item In contrast to SN\,2011fe, $(B - V)_0$ is $\sim$0.5 mag bluer for \abc\ $\sim$2\,d after $t_0$. Furthermore, the $(B - V)_0$ colors of \abc\ show no evolution over the first $\sim$7\,d  of observations.

    \item Finally, we show that the early light curve evolution and colors of \abc\ are well matched by the theoretical models of \citet{2016ApJ...826...96P} that include significant mixing of $^{56}$Ni into the outer layers of the SN ejecta.

\end{itemize}
%
Taken together, these observations all indicate that the nucleosynthetic products from the explosion are well mixed throughout the SN ejecta. There are elements of the PDD models from \citet{2014MNRAS.441..532D} that are attractive for explaining \abc, in particular, the strong \ion{C}{2} absorption seen at early times. In the future, it would be useful to investigate more detailed PDD models that incorporate strong $^{56}$Ni mixing to see if they better replicate the observations of \abc\ as it is otherwise difficult to distinguish between these two scenarios \citep{2017MNRAS.472.2787N}.\footnote{Both \citet{2014MNRAS.441..532D} and \citet{2016ApJ...826...96P} consider models with strong mixing \textit{and} interaction, but those studies do not replicate each of the salient features of \abc.}

Extremely early observations of young SNe provide a ``smoking
gun'' to probe the mixing level in the ejecta, which, in turn, is 
a result of the explosion mechanism. Wide-field, high-cadence surveys, such as the Zwicky Transient Facility \citep{2016PASP..128h4501B} and ATLAS \citep{2011PASP..123...58T,2013RSPTA.37120269T}, will discover a large number of very young SNe over the next few years, allowing us to extend our studies beyond single objects. While the sample of extremely young SNe Ia will grow by more than an order of magnitude, the detection of shock breakout cooling and ejecta-companion interaction will prove challenging. Given the diminutive size of WDs, the thermal emission following shock breakout can only be detected to $\sim 10\,\mathrm{Mpc}$ on 1-m class telescopes. Furthermore, only $\sim$10\% of single-degenerate progenitors are expected to give rise to detectable emission following the collision of the SN ejecta with the binary companion \citep{2010ApJ...708.1025K}. Despite these limitations, this study of \abc\ shows that the early detection of SNe Ia can probe explosion physics by measuring the amount of mixing in the SN ejecta. Moving forward, a large sample of such objects will enable strict constraints on the proposed explosion mechanisms for Type Ia SNe.

Finally, we close by emphasizing the importance of fast-response photometric and spectroscopic follow-up campaigns. Without the early recognition of the youth of this SN, and the associated follow-up, much of the analysis presented herein would not have been possible. The ability to trigger such observations is essential to improve our understanding of the physics of SNe Ia.

@@

AAM is funded by the Large Synoptic Survey Telescope Corporation in support of the Data Science Fellowship Program. YC thanks support from a postdoc fellowship in the eScience institute, University of Washington.

@@

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network. These results made use of the Discovery Channel Telescope at Lowell Observatory. Lowell is a private, non-profit institution dedicated to astrophysical research and public appreciation of astronomy and operates the DCT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University and Yale University.  The upgrade of the DeVeny optical spectrograph has been funded by a generous grant from John and Ginger Giovale. Based on observations made with the Nordic Optical Telescope, operated by the Nordic Optical Telescope Scientific Association at the Observatorio del Roque de los Muchachos, La Palma, Spain, of the Instituto de Astrofisica de Canarias.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{lightcurve.pdf}
  \caption{UV, optical, and NIR light curves for \abc. 
  \textit{Top}: $BVgri$ light curves from the P48, P60, and LCO-1m 
  telescopes. The solid lines represent the best-fit model from 
  SALT2 (see \ref{sec:classification}).
  \textit{Bottom}: UV and NIR light curves from the \textit{Swift} 
  and SPM-1.5m telescopes, respectively.
  For both panels each light curve is represented with a different 
  color, while the different symbols correspond to different 
  instruments, as detailed in the legends. The legends also list 
  offsets applied to each light curve. Photometry is shown in the 
  AB mag system, with the exception of the $BV$ bands, which are 
  shown in the Vega system. The photometry shown has not been 
  corrected for line-of-sight extinction.
  }
  \label{fig:ugly_LC}
\end{figure}

>>>>

\begin{itemize}

    \item Extrapolation of the early light curve shows that the initial
    detection of \abc\ occurred only $0.15\pm_{0.07}^{0.15}$~d after the time
    of first light, $t_0$.

@@

    \item Assuming that $v_\mathrm{ph} \propto t^{-0.22}$, then
    $t_\mathrm{exp} \approx t_0$. A short dark phase, as this implies, is
    likely the result of either strong $^{56}$Ni mixing or interaction of the
    SN ejecta with \textbf{nearby, unbound} material.

@@

    \item In contrast to SN\,2011fe, $(B - V)_0$ is $\sim$0.5 mag bluer for
    \abc\ at $t \approx -16$\,d. Furthermore, the $(B - V)_0$ colors of \abc\
    show no evolution over the first $\sim$7\,d of observations.

    \item Finally, we show that the early light curve evolution and colors of
    \abc\ are best \textbf{matched by the pulsational delayed detonation
    models of \citet{2014MNRAS.441..532D}} and the ejecta-mixing models of
    \citet{2016ApJ...826...96P}.

\end{itemize}
%
Taken together, these observations \textbf{suggest that the early emission
from \abc\ is due to either the collision of the SN with nearby, unbound
material and/or that there is significant mixing of $^{56}$Ni in the SN
ejecta.} The PDD models from \citet{2014MNRAS.441..532D} are particularly
attractive for explaining \abc, \textbf{because they produce} strong
\ion{C}{2} absorption at early times. In the future, it would be useful to
investigate more detailed PDD models that incorporate strong $^{56}$Ni mixing
to see if they better replicate the observations of \abc\ as it is otherwise
difficult to distinguish between these two scenarios.

Extremely early observations of young SNe provide a ``smoking
gun'' to probe the mixing level in the ejecta, which, in turn, is 
a result of the explosion mechanism. Wide-field, high-cadence surveys, such as the Zwicky Transient Facility \citep{2016PASP..128h4501B} and ATLAS \citep{2011PASP..123...58T,2013RSPTA.37120269T}, will discover a large number of very young SNe over the next few years, allowing us to extend our studies beyond single objects. While the sample of extremely young SNe Ia will grow by more than an order of magnitude, the detection of shock breakout cooling and ejecta-companion interaction will prove challenging. Given the diminutive size of WDs, the thermal emission following shock breakout can only be detected to $\sim 10\,\mathrm{Mpc}$ on 1-m class telescopes. Furthermore, only $\sim$10\% of single-degenerate progenitors are expected to give rise to detectable emission following the collision of the SN ejecta with the binary companion \citep{2010ApJ...708.1025K}. Despite these limitations, this study of \abc\ shows that the early detection of SNe Ia \textbf{may} probe explosion physics by measuring the amount of mixing in the SN ejecta. Moving forward, a large sample of such objects will enable strict constraints on the proposed explosion mechanisms for Type Ia SNe.

Finally, we close by emphasizing the importance of fast-response photometric and spectroscopic follow-up campaigns. Without the early recognition of the youth of this SN, and the associated follow-up, much of the analysis presented herein would not have been possible. The ability to trigger such observations is essential to improve our \textbf{physical} understanding of Type Ia SNe.

@@

AAM is funded by the Large Synoptic Survey Telescope Corporation in support of
the Data Science Fellowship Program. YC thanks support from a postdocal
fellowship at the eScience Institute, University of Washington.

@@

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network. These results made use of the Discovery Channel Telescope at Lowell Observatory. Lowell is a private, non-profit institution dedicated to astrophysical research and public appreciation of astronomy and operates the DCT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University and Yale University. The upgrade of the DeVeny optical spectrograph has been funded by a generous grant from John and Ginger Giovale. Based on observations made with the Nordic Optical Telescope, operated by the Nordic Optical Telescope Scientific Association at the Observatorio del Roque de los Muchachos, La Palma, Spain, of the Instituto de Astrof\'isica de Canarias.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{lightcurve.pdf}
  \caption{
  %
  UV, optical, and NIR light curves for \abc. \textit{Top}: $BVgri$ light
  curves from the P48, P60, LCO-1m, and \textit{Swift} telescopes. The solid
  lines represent the best-fit model from SALT2 (see
  \S\ref{sec:classification}). \textit{Bottom}: UV and NIR light curves from
  the \textit{Swift} and SPM-1.5m telescopes, respectively. For both panels
  each light curve is represented with a different color, while the different
  symbols correspond to different instruments, as detailed in the legends. The
  legends also list offsets applied to each light curve. Photometry is shown
  in the AB mag system, with the exception of the $BV$ bands, which are shown
  in the Vega system. These \textbf{light curves have} not been corrected for
  line-of-sight extinction.}
  %
  \label{fig:ugly_LC}
\end{figure}
